---
title: "Gapminder Data Exploration"
authors: "Erin Roberts,Melissa Hoffman,Maisaa Albkerat"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(curl)
library(ggplot2)
library(plyr)
gapminder_location<-curl(url = "https://raw.githubusercontent.com/resbaz/r-novice-gapminder-files/master/data/gapminder-FiveYearData.csv")
gapminder<-read.csv(gapminder_location)
knitr::opts_chunk$set(echo = FALSE, fig.width = 5, fig.height = 5, warning = FALSE, message = FALSE)
```

\begin{center} 
Erin Roberts, Melissa Hoffman, Maisaa Albkerat
\end{center}
\vspace{12pt}

## Life Expectancy Trends
\vspace{12pt}

```{R Life exp, echo=FALSE,fig.width=5.5,fig.height=3}
ggplot(data=gapminder, aes(x=year, y=lifeExp))+
  geom_point()+geom_smooth(method="lm")
```

```{R "lifeExp_time", echo=FALSE, include=FALSE}
LifeExpstats <- lm(lifeExp ~ year, data=gapminder)
LifeExpstats
summary(LifeExpstats)
```

Figure 1. Shows a positive correlation between global life expectancy over time from 1952-2007. There is a slight positive correlation, R-squared = `r summary(LifeExpstats)$r.squared`, and the correlation is significant, P-value = 2.2e-16 `r summary(LifeExpstats)$p.value`.  
\vspace{20pt}


``````{R Life exp for cont, echo=FALSE,fig.width=6.7,fig.height=3}
ggplot(data=gapminder, aes(x=year, y=lifeExp, color=continent))+
  geom_point()+geom_smooth(method=lm)
```

Figure 2. Shows life expectancy for 1952-2007, but this time the data is color-coded by continent. Additionally, a trendline allows us to see that life expectancy has increased in all continents over time.


## Life Expectancy and Population
\vspace{12pt}

```{r Life exp with pop growth by country, echo=FALSE, figure.width=6.7,fig.height=3}
ggplot(data=gapminder,aes(x=pop, y=lifeExp, by=country, color=continent)) + geom_line() + geom_point()

```

Figure 3. This graph depicts how population growth for each country affects life expectancy. There is a strongly positive correlation between population growth and life expectancy at low population levels. The slope levels off as the population grows however. Asia has larger population levels and still shows an overall positive relationship between population growth and life expectancy for population sizes above 5 x 10^8.
\vspace{20pt}

```{r population per year per country, echo=FALSE, figure.width=6.7,fig.height=3}
ggplot(data=gapminder, aes(x=year, y=pop, by=country, color=continent)) + geom_line() + geom_point()
```

Figure 4.This graph of population over time shows a positive correlation in between 1952 and 2007, though Africa appears to be the most stagnant in terms of growth. The R-squared value is `r summary(LifeExpstats)$r.squared`. Asia has the highest population levels of any country studied. 

  

```{r gdpPercap_year,echo="FALSE",include="FALSE"}
reg <- lm(gdpPercap ~ year, data=gapminder)
reg
summary(reg)
```

##Relationship between gdpPercap and year
\vspace{12pt}

```{r scatterplot, fig.width=5.5, fig.height=3, echo=FALSE}
ggplot(data = gapminder, aes(x = year, y = gdpPercap)) +
  geom_point() +  scale_y_log10() + geom_smooth(method="lm")

```

Figure 5. There is significant positive correlation between gdpperCap and year, it is increasing by years but different by countries and continents.
R-squared= `r summary(reg)$r.squared`. 
\vspace{20pt}


```{r  fig.width=6.7, fig.height=3, echo=FALSE}
ggplot(data = gapminder, aes(x=year, y=gdpPercap, color=continent)) +
   geom_point()+  scale_y_log10()+ geom_smooth(method="lm")

```

Figure 6. There is a positive correlation between GdpPercap and year for all continents.


