---
title: "Readme.md"
authors: "Erin Roberts, Melissa Hoffman, Maisaa Albkerat"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(curl)
library(ggplot2)
library(plyr)
gapminder_location<-curl(url = "https://raw.githubusercontent.com/resbaz/r-novice-gapminder-files/master/data/gapminder-FiveYearData.csv")
gapminder<-read.csv(gapminder_location)
knitr::opts_chunk$set(echo = FALSE, fig.width = 5, fig.height = 5, warning = FALSE, message = FALSE)
```

## Gapminder Data Exploration


```{R Life exp, echo=FALSE,fig.width=5.5,fig.height=3}
ggplot(data=gapminder, aes(x=year, y=lifeExp))+
  geom_point()
```

Figure 1. Shows life expectancy from 1952-2007 for all continents. It is difficult to see any trends in this figure. 


``````{R Life exp for cont, echo=FALSE,fig.width=6.5,fig.height=3}
ggplot(data=gapminder, aes(x=year, y=lifeExp, color=continent))+
  geom_point()+geom_smooth(method=lm)
```

Figure 2. Shows life expectancy for 1952-2007, but this time the data is color-coded by continent. Additionally, a trendline allows us to see that life expectancy has increased in all continents over time.

```{r Life exp with pop growth by country, echo=FALSE, figure.width=5.5,fig.height=3}
ggplot(data=gapminder,aes(x=pop, y=lifeExp, by=country, color=continent)) + geom_line() + geom_point()

```

Figure 3. This graph depicts how population growth for each country affects life expectancy.

```{r population per year per country, echo=FALSE, figure.width=6.5,fig.height=4}
ggplot(data=gapminder, aes(x=year, y=pop, by=country, color=continent)) + geom_line() + geom_point()
```

Figure 4. This graph illustrates how population size for each listed country has changed every year.


